---
title: "Regression Model"
author: "Luiz Bergo"
date: "24/11/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(car)
```

## Summary

This is a project assignment as per request to complete the Regression Model course from Johns Hopkins University powered by Coursera. The main idea is to analyse the ***mtcars*** dataset provided in R Packages and answer the question that raised:

 **“Is an automatic or manual transmission better for MPG?”**
 
 Also is asked to  "Quantify the MPG difference between automatic and manual transmissions".

At the following ssesions the data is discussed to properly answer the question.


## Exploratory Data Analysis

The **mtcars** dataset is composed for 32 samples and 11 variables. 

```{r data}
data("mtcars")
head(mtcars)
dim(mtcars)

```

The idea is to predict the influence of type of transmission (automatic or manual) in the fuel consumption **mpg** (Miles/(US) gallon). The remaining variable are descripted as follow:

**cyl**	- Number of cylinders  
**disp** -	Displacement (cu.in.)  
**hp** - 	Gross horsepower  
**drat** -	Rear axle ratio  
**wt**	- Weight (1000 lbs)  
**qsec** - 1/4 mile time  
**vs**	- Engine (0 = V-shaped, 1 = straight)  
**am**	- Transmission (0 = automatic, 1 = manual)  
**gear**	- Number of forward gears  
**carb**	- Number of carburetors  

Correlation of outcome with possible predictors:

```{r}
cor(mtcars)[1,]
```

A comparison between the variables are depicted in the **Appendix**.

## Model Adjustment

The first idea is to evaluate fuel consumption acording to the transmisison.

```{r transmission_plo}
cars <- mtcars
cars$am <- as.factor(cars$am)
levels(cars$am) <- c("Automatic", "Manual")

boxplot(mpg~am, data = cars, main = "Fuel consumption for type of transmission",
        col=c("red","blue"), xlab = "Transmission",ylab = "Consumption - (Miles/(US) gallon)")


fit0 <- lm(mpg ~ factor(am), data=cars)
tcritical <- qt(0.975, fit0$df)
print(paste("Critical t value at 95% of confidence to reject null hypothesis :", round(tcritical, 3)))
summary(fit0)
```


The t-values for coeficients are greater than $t_c$ so it is validy assume some correlation between the variables and transmission explain about **34%** of variance at fuel comsuption. This lead us to interpret that **manual transmission cars** tend to consume less than **automatic**.

Considering that only **34%** of variation is explained for this model some other variables may be included for a better understandig. Such variation can either be observed in the plot of residuals.

```{r residual_fit0}
plot(fit0$residuals, main = "Residual plot from the first model")
```

The first approach is to evaluate a linear fit including all the variables available as predictors.

```{r fit_all}
fitall <- lm(mpg ~ ., data=cars)
summary(fitall)
plot(fitall$residuals, main = "Residual plot for complete model")
```

The model including all variables change has a better result than the first model with just **am** (transmition) as a predictor with about **81%** of variance explained.

Although it is possible to observe a high variation inflation, suggesting the model can be simplified.

```{r}
vif(fitall)
```

Observing the plot and correlation between variables depicted at the *Appendix* it is possible to propose new fits observing that onde more correlated to mpg.

```{r}
fit1 <- lm(mpg ~ factor(am) + wt, data=cars)
vif(fit1)
summary(fit1)$adj.r.squared
```
With this model **74%** of variance is explained.

To improve the model the **hp** predictor was included and the variance explained is **82%**, better than the model with all the predictors.

```{r}
fit2 <- lm(mpg ~ factor(am) + wt + hp, data=cars)
vif(fit2)
summary(fit2)$adj.r.squared
```

Other predictors were tested aiming increase in the adjusted R square factor, and the final model includes **am**, **wt**, **hp** and **cyl** as predictors and the variance explained is **83%**.

A caveat is the variance inflation observed in the **hp** predictor since it is very correlated to **cyl** predictor.

```{r}
fit3 <- lm(mpg ~ factor(am) + wt + hp + cyl, data=cars)
vif(fit3)
summary(fit3)$adj.r.squared
```

It´s worthy to note the residual plot is quite near from that obtained using all the predictors.


```{r}
plot(fit3$residuals, main = "Residual plot for final model")

par(mfrow = c(2,2))
plot(fit3)
```


The coeficients for the final model are:

```{r}
summary(fit3)$coef
```

## Conclusion

The confidence interval for the coeeficients are:

```{r}
confint(fit3)
```

Suposing the same values for predictors **wt**, **hp** and **cyl**, manual cars are at average **4%** more economic than automatic cars.

Based on the final model it´s possible to conclude that cars with manual transmition are slightly economic than automatic car, so they are a better regarding this aspect. 

## Apendix

```{r}

pairs(cars)
```

